Marine species do not need visas: the transboundary nature of the world’s exploited marine species

### 2020-08-06 Scripts/Manuscript_Draft.Rmd
Incorporated co-authors comments

---
title: "Marine species do not need visas: the transboundary nature of the world’s exploited marine species"
author: "Authors: Juliano Palacios-Abrantes*^1^, Gabriel Reygondeau^1,2^, Colette Wabnitz^1,3^ and William W.L. Cheung^1^"
date: "Affiliations: ^1^Institute for the Oceans and Fisheries, University of British Columbia, Vancouver, Canada. ^2^Yale University, New Haven, Connecticut, USA. ^3^Stockholm Resilience Center, Stockholm University, Stockholm, Sweden."
subtitle: "CORECTIONS ARE IN BOLD. DO NOT UNBOLD"
# subtitle: "Most neighbor nations share marine exploited species"
output:
  word_document:
    # reference_docx: ./Manuscript_Instructions/science_template.docx
  pdf_document:
    fig_caption: yes
geometry: margin=1in
editor_options:
  chunk_output_type: inline
bibliography: ../References/FishForVisa_Ref.bib
csl: ../References/nature_cc.csl
---

@@

Regulatory boundaries and species distributions are often not aligned, especially in the case of marine species crossing multiple Economic Exclusive Zones (EEZs). This represents a challenge for fisheries management as policies tend to focus at the national level, yet collaborations are needed to maximize long-term ecological, social and economic benefits. Here, we combined species distributions and the spatial delimitation of EEZs to identify the number of commercially exploited marine species that are shared between neighboring nations. We found that 67% of the world's identified fished species are transboundary (n = 633) and were responsible for 82% of the global-EEZs catches (48 million tons) and 81% of fisheries revenue (USD 77 billion) in the last decade. Some countries received over 90% of their catch and economic benefits from a few shared resources. Our analysis suggests that fisheries targeting transboundary species appear **more intensively fished** than those mainly catching non-transboundary species. Our study has direct implications **for fisheries management** of transboundary species, pointing to the urgent need for strengthened international cooperation in the management of these fisheries.

@@

Distributions of marine species around the world are not constrained by human-made boundaries; rather they are shaped by biotic and abiotic factors defined by their evolutionary history [@Hutchinson:1957nt; @Nelson:2016kk]. Thus, a species can have a broader distribution (cosmopolitan) or be endemic [@Nelson:2016kk]. However, fisheries management is predicated on the definition of “stocks” that, in most cases, are delineated by human-made spatial boundaries that often do not correspond to biologically-meaningful population units [@Song:2017iua; @FredstonHermann:2018kp]. The delineation of Economic Exclusive Zones (EEZs) under Part V of the United Nations Convention on the Law of the Seas (UNCLOS) in the early 80s, for instance, extended political boundaries **from 12 (territorial see) to 200 nautical miles** across the distribution of many species, creating shared stocks between nations [@UnitedNations:1986tl; @osthagen:2020hd]. **Shared stocks can be classified in three non-exclusive categories; transboundary, that is stocks that cross the EEZs of two or more coastal states; straddling, that cross neighboring EEZs and the high seas; and highly migratory stocks, that cross non-neighboring EEZs and the high seas (mainly tunas) [@Munro:2004th]. For the purpose of this research we focus on identifying the transboundary nature of marine fished species, that is, those shared by neighboring EEZs, noting that in some cases these might also be present in the high seas (e.g. straddling).** Theories and empirical experience have shown that fisheries targeting resources that straddle between political boundaries complicate fisheries management and potentially reduce its effectiveness to achieve stated objectives [@Munro:2002uf; **@Englander:2019dq**]. For instance, climate-driven changes in species distributions have led to conflict between nations due to changes in the proportion of captures [@Munro:2002uf], quota allocation, and fishery newcomers [@Spijkers:2017ij]. Moreover, climate change is likely to exacerbate these types of conflict and presents new challenges for political relations between neighboring countries [@Pinsky:2018cb] and for fisheries management [@Miller:2013iv].

Forty years after the formal adoption of UNCLOS **and the consequent 1995 United Nations Fish Stock Agreement for the cooperation on the management and conservation of straddling resources [@UnitedNations:1995fs], accurate estimates of the number of exploited marine species that are shared by neighboring nations is still unclear. Early data limitations on species' distributions allowed only for an informed guess of 500 to 1500 transboundary stocks [@Caddy:1997ue] while a more recent study explored the historical contribution of 344 shared taxa following a literature review [@Teh:2015gd].** Thus, no method has included the actual distribution of species in the quantification of transboundary stocks. Here, we aimed to estimate the number of exploited marine species shared by neighboring countries and determine their contribution to global and regional catch and fisheries revenue. Moreover, we categorize species according to their exploitation rate while identifying differences by habitat preference. We hypothesize that the methodological constraints of previous work has resulted in an underestimation of the contribution of these species to global catch and revenue, and that the proportion of transboundary species is larger than previously estimated.

We overlaid the known distribution of 938 commercially valuable marine species responsible for an average of 96.5% of global EEZ catch reconstructed by the *Sea Around Us* at the species level between 2005 and 2014, and 280 EEZs of 198 coastal **countries (See methods**). We define a ‘stock’ unit as a species in an EEZ, instead of using a population ecology definition (genetically or morphological distinct units) due to lack of such biological information for almost all the species included in this study [@Teh:2015gd]. While we acknowledge that a species could have multiple stocks within an EEZ, many fisheries within a country or EEZ are managed at the species instead of stock level **(or even as a group of species) like the case of shrimps (e.g. Litopenaeus stylirostris or Farfantepenaeus californiensis) in the Pacific coast of Mexico [@DiarioOficialdelaFederacionDOF:2018wt], hammerhead shark (*Sphyrna zygaena*) in Peru [@MAP:2017uf].** Also, recent research has suggested strong connectivity between stocks, even when separated by thousands of kilometers [@Ramesh:2019va] providing additional ecological ground for our analysis **while still making it relevant from a management perspective**. For this analysis we only considered shared species between neighboring EEZs, rather than the species’ extended distribution (e.g. we did not include the high seas). We rely on multiple data sources including occurrence, distribution models and catch data and only consider a species to be present in a grid cell if all data sources predicted positive occurrence (see Methods).

@@

**We identified 633 transboundary exploited species in the world (67.5% of the 938 species analyzed), almost double the previous estimates [@Teh:2015gd].** Between 2005 and 2014, national fleets targeting transboundary species within EEZs **caught** an **annual** average of 48.5 million tons, representing 82.3% of the EEZ-derived catches reconstructed by the *Sea Around us* at the species level, and generating **a yearly average of USD 77,591 million in fishing revenue during the same time period**. This is equivalent to 81.3% of the global fishing revenue of the same period. These values are more than the 34.2 million tons and USD 40,187 million (**converted to in 2019 USD**) in fishing revenue from shared stocks previously estimated [@Teh:2015gd]. **Such difference could be due to the utilization of reported catch data by Teh and Sumaila instead of reconstructed data and a time period difference (Teh and Sumaila's study ends in 2006). To compare the two studies, we used the FAO global reported data identified at the species level by the *Sea Around Us* and found that in 2006, the 633 transboundary species accounted for 40.7 million tons. While this is not necessarily a direct comparison (e.g. methodology differences might still exist), it suggests that the participation of transboundary species in global catch and fishing revenue might have been previously underestimated due to an incomplete knowledge on the transboundary nature of marine fished species**

Based on our assessment (see Methods), 582 transboundary species were also identified as "discrete" species, meaning that the same species could be shared by two neighboring EEZs (transboundary) but not with a third neighbor (discrete). For example, our study shows that Peruvian anchoveta (*Engraulis ringens*) is fished by Peru, Chile, and Ecuador but only transboundary between Peru and Chile and thus discrete for Ecuador (**spatial threshold between Ecuador and Peru < 10%**) **[@Cashion:2018cg]**. The **annual average contribution of discrete species (in our previous example would be only considering Ecuador's anchoveta catch) between 2005 and 2014** was 7.5 million tons representing USD 12,941 million in fishing revenue **over the same time period**. Within these species, 305 were considered exclusively discrete, that is, they could be fished by neighboring nations but were not considered transboundary **according to our methods of determination (See methods)**. These species represented only **an annual average of** 2.8 million tons and 4,282 million USD between 2005 and 2014. These results underscore the importance of transboundary species at the global level.

We group countries according to the United Nations sub-regions and found that transboundary species are particularly economically important for Northern America (**average country** revenue = 4,680 $\pm$ 6,000 million USD) and Eastern Asia (3,779 $\pm$ 3,093 million USD). The estimated **per country** revenues from transboundary species in these two regions are significantly different from other regions (one-way ANOVA; *DF*(16,165) = 5.081, *p* < 0.001, $\alpha$ = 0.05) (Fig. 1B). China (USD 7,284 million) in **Eastern** Asia and the USA **in Northern America** (USD 11,604 million) contribute 55% and 82% to the **annual average** revenues. In addition to China and the USA, Russia (USD 7,379 million), Peru (USD 6,044 **million**) and Japan (USD 3,907 **million**) are among the top five countries with the most fishing revenue generated from transboundary species in the last decade. These five nations are responsible for 41% of the global fisheries revenue from transboundary species.

@@

![**Figure 1.** Number of transboundary species and their contribution to global fisheries catches and revenue. A) The number of species and amount of revenue are represented by color coding of EEZs and land polygons, respectively. B) Contribution of transboundary species to regional revenue (left) and catch (right). Regions according to the United Nations sub-regions. Points = mean $\pm$ sd. Revenue in 2010 real USD.](../Figures/Fig1.png)

Peru and Russia, responsible for the two largest fisheries in the world [@FAO:2018KK], lead transboundary catch statistics harvesting over 5.8 million tons annually each on average between 2005 and 2014 (Supplementary Fig. S1). Peru's catches are mainly composed of Peruvian anchoveta (*Engraulis ringens*), shared with Chile, and account for 79% of the **national** transboundary species production. Recently these two countries have signed an agreement to work towards standardized stock assessments through coordinated management of this species [@UNDP:2016wp]. **Similarly, Russia, Japan and the USA celebrate a management agreement over Alaskan pollock (*Theragra chalcogramma*) in the Bering Sea since 1988 [@NOAA:gX6CtUTk].** Transboundary species also contribute largely to fisheries in **Eastern** Asia (one-way ANOVA; *Df*(16,158) = 2.265, *p* = 0.005, $\alpha$ = 0.05). Particularly, China the world's top fish producer**,** [@FAO:2018KK] **gets** one third (5.1 million tons) of the total catches from transboundary species, followed by Japan (1.8 million tons) and South Korea (1.06 million tons). Differences in transboundary fisheries can also be illustrated in terms of catch-revenue/EEZ $km^2$ (Fig. 2). As an example, the aggregated area of all EEZs of Northern Europe is 3.3 million $Km^2$, the $6^{th}$ smallest of the 17 groups analyzed (Supplementary Table S1), however, it has the second **highest average** revenue (USD 246.7 thousand per $Km^2$) and catch (260.1 tons per $Km^2$) of transboundary species per EEZ area between 2005 and 2014, but only significantly differing from Sub-Saharan African and Southern Europe countries (*p* < 0.05).

@@

![Figure 2. Weighted benefits of transboundary species by Km2 and UN sub-region. A) Units in thousand 2019 USD. B) Units in tons. Points are sub-region averages by country $\pm$ s.d.](../Figures/Fig2.png)

**We determined the exploitation category of each species according to the catch trend within each EEZ (Fig. 3). While previous studies demonstrated that catch trends can be used to infer whether a stock is healthy, re-building, over-exploited or collapsed [@Kleisner:2011wn], several factors can influence stock status. Therefore, our intention here is not to infer stock status but to show the catch trend of transboundary species between 2005 and 2010 relative to historic catch since 1951 (See methods).** We found significant differences in exploitation categories based on whether or not a species was transboundary (MANOVA, *Df*(2,459) = 29.406, *p* < 0.001). Specifically, **occasions where a transboundary species is only fished by one country are generally less likely to have a decreasing catch trend compared to when the species is shared by neighboring countries (Supplementary Table S2)**. Significant differences were found for all categories within transboundary species (one-way ANOVA, *Df*(2,459) = 53.93, *p* $\leq$ 0.001, $\alpha$ = 0.05; Tukey's post hoc test *p* < 0.05) but only Category C for discrete species (one-way ANOVA, *Df*(2,382) = 25.58, *p* < 0.001; $\alpha$ = 0.05; Tukey's post hoc test for Categories C-A and C-B *p* < 0.001, and B-A 0.058).

```{r Call_Circ_Country, eval = F, fig.width=6, fig.height=4, echo=FALSE, fig.pos= "H", fig.cap="Number of transboundary species by catch category. Category A, increasing catch; Category B, stable catch; Category C, decreasing catch"}

@@

![Figure 3. Number of transboundary species by catch trend and EEZ. Category A, Growing trend; Category B, Stable trend; Category C, Shrinking trend](../Figures/Fig3.png)

@@

Transboundary fisheries are abundant and of economic importance for some of the top countries with past fisheries-related conflicts including Canada, the USA, Russia and the European Union (EU) [@Spijkers:2019cz]. For example, since 2007, the EU, Norway, Iceland, and the Faroe Islands have been at odds over the size and relative allocation of total allowable catches for Atlantic mackerel (*Scomber scombrus*) due to a climate-driven distribution shift [@Spijkers:2017ij]. Mackerel is a transboundary species that **straddles into the high seas** that contributed an average of 598.4 thousand tons to **total** catches (8.18% of total) and USD 986.24 million in **fishing** revenue (7.16% of total) to the countries involved in the dispute between 2005 and 2014. The relatively small contribution of mackerel to the involved countries suggests that even a small fraction of the year's catch can lead to important conflicts between neighboring countries. Climate change is expected to continue changing the distribution and share proportion of fish stocks among countries up to a degree that new transboundary species will emerge [@Pinsky:2018cb], while some species will exit EEZs [@Oremus:2020id]. Exploring the detailed effects of climate change in the distribution of shared species is key to develop local adaptation methods to anticipate negative impacts to sustainability. **For example, understanding how climate change will modify the proportion that each neighboring EEZ has over a sharing species [@PalaciosAbrantes:2020vt], when this change is going to happen, and the economic consequence of such effects can make international treaties more resilient to climate change [@RashidSumaila:tp].**

Over 130 marine fish species occur in tropical and subtropical waters around the world [@Nelson:2016kk], from highly migratory species associated to pelagic-oceanic ecosystems like tunas (*Thunnus sp.*), to less mobile reef-associated species like greater amberjack (*Seriola dumerili*), to species found in demersal ecosystems like gilthead seabream (*Sparus aurata*). Species associated to pelagic-oceanic ecosystems are the only group whose EEZ range (i.e., the number of EEZs where the species occur as transboundary) was significantly different from other groups (one-way ANOVA; *DF*(5,597) = 53.82, *p* < 0.001, $\alpha$ = 0.05; Tukey's post hoc test *p* < 0.05) with a median of 40 EEZs per species. The median for species of all other ecosystem preferences is close to, or less than, 20, as many of these species have a narrower distribution or are less mobile (Fig. 4A). **[Colette] This result is likely an effect of the broad distribution that many of the species with preference for pelagic-oceanic ecosystems have. Many of these species are actually highly migratory straddling species (species that go from one EEZ to another while crossing the high seas). Thus, the number of non-neighboring EEZs sharing a highly migratory species can be over 100, like the case of Bigeye tuna (Thunnus obesus) (Fig. 4A). Due to their large migration patterns and presence in areas beyond national jurisdiction, many highly migratory species are managed by Regional Fisheries Management Organizations (RFMOs).**

CW. because they are migratory they fall under the management of RFMOs...

@@

![Figure 4. Number of EEZs shared by transboundary species. A) Number of EEZs by species organized by associated habitat according to FishBase. Showing only species that share > 20 EEZs. B) Total catch between 2005 and 2014 of top five fishing countries capturing the top five shared species of each ecosystem preference (shown in bold on A). Note that there could be > 5 species due to similar sharing values. **Other consist of species that have no classification according to FishBase.**](../Figures/Fig4.png)

**However, this category also includes species that are not considered highly migratory but still shared by numerous neighboring EEZs (e.g. Belone belone). Species with narrower ranges are often managed at the national level due to localized populations [@Payne:2008vm] but the potential boundary-crossing of these species must not be ignored. In addition, the world of fisheries is highly connected through larvae dispersal and so, even centralized populations could be supplying populations thousands of kilometers away [@Ramesh:2019va]. Despite coral reef-related species being among the less shared neighboring EEZs identified in this study (Fig. 4), coral regions often share multiple species through larvae connectivity [@Schill:2015fh] and adult movement [@Levin:2018in]. In some areas like the Caribbean and the Western Indian Ocean, transboundary marine protected areas have been identified as potential tools for achieving fisheries and conservation goals [@Levin:2018in; @Perez:2019dm].** The proper management of coral reef species is very important for many coastal communities, particularly in developing regions, that are highly dependent on subsistence fishing for food security and livelihoods [@CisnerosMontemayor:2016gq; @Wabnitz:2018gf]. For example, all of the top five countries fishing transboundary reef and pelagic-oceanic associated species (Fig. **4B**) are among the highest fish consumption countries [@FAO:2018KK]. In these countries, both pelagic and reef fishes contribute substantially to both local food security and livelihood [@Cabral:2018hc].

@@

Occurrence data was collected by Reygondeau (2019) from five publicly available repositories: FishBase (http://fishbase.org), the Global Biodiversity Information Facility (GBIF; https://www.gbif.org/), the Ocean Biogeographic Information System (OBIS; https://obis.org/), the Intergovernmental Oceanographic Commission (IOC; http://ioc-unesco.org), and the International Union for Conservation of Nature (IUCN; https://www.iucn.org/technical-documents/spatial-data) [@Reygondeau:2019uh].

@@

Categories <- c("A - Growing Fishery",
                  "B - Stable Fishery",
                  "C - Shrinking Fishery",
                  "No Status")

>>>>

---
title: "Marine species do not need visas: the transboundary nature of the world’s exploited marine species"
author: "Authors: Juliano Palacios-Abrantes*^1^, Gabriel Reygondeau^1,2^, Colette Wabnitz^1,3^ and William W.L. Cheung^1^"
date: "Affiliations: ^1^Institute for the Oceans and Fisheries, University of British Columbia, Vancouver, Canada. ^2^Yale University, New Haven, Connecticut, USA. ^3^Stockholm Resilience Center, Stockholm University, Stockholm, Sweden."
# subtitle: "Most neighbor nations share marine exploited species"
output:
  word_document:
    # reference_docx: ./Manuscript_Instructions/science_template.docx
  pdf_document:
    fig_caption: yes
geometry: margin=1in
editor_options:
  chunk_output_type: inline
bibliography: ../References/FishForVisa_Ref.bib
csl: ../References/nature_cc.csl
---

@@

Regulatory boundaries and species distributions are often not aligned, especially in the case of marine species crossing multiple Exclusive Economic Zones (EEZs). This represents a challenge for fisheries management as policies tend to focus at the national level, yet collaborations are needed to maximize long-term ecological, social and economic benefits of shared marine species. Here, we combined species distributions and the spatial delineation of EEZs to identify the number of commercially exploited marine species that are shared between neighboring nations. We found that 67% of the world's identified fished species are transboundary (n = 633). Between 2005 and 2014, fisheries targeting these species within the global-EEZs caught in average 48 million tonnes representing an average of USD 77 billion in fishing revenue per year. Some countries received over 90% of their catch and economic benefits from a few shared resources. Our analysis suggests that catches from transboundary species appear **to be declining** more than those from non-transboundary species. Our study has direct implications **for managing** fisheries of transboundary species, highlighting the need for strengthened international cooperation.

@@

Distributions of marine species around the world are not constrained by human-made boundaries; rather they are shaped by biotic and abiotic factors as well as the species' evolutionary history [@Hutchinson:1957nt; @Nelson:2016kk]. A species can be widely distributed (cosmopolitan) or endemic [@Nelson:2016kk]. However, fisheries management is predicated on the definition of “stocks” that, in most cases, are delineated by human-made spatial boundaries while such delineation often does not correspond to biologically-meaningful population units [@Song:2017iua; @FredstonHermann:2018kp]. The delineation of Exclusive Economic Zones (EEZs) under Part V of the United Nations Convention on the Law of the Seas (UNCLOS) in the early 80s, for instance, extended political boundaries **from 12 to 200 nautical miles.** The boundaries of EEZs cut across the distribution of many species, creating shared stocks between nations [@UnitedNations:1986tl; @osthagen:2020hd]. **Shared stocks can be classified as three non-exclusive categories; "transboundary" - stocks that cross the EEZs of two or more coastal states; "straddling" - stocks that, cross neighboring EEZs and the adjacent high seas; and "highly migratory" - stocks that cross non-neighboring EEZs and the high seas (mainly tunas) [@Munro:2004th]. Here, our study focuses on the "transboundary" and "straddling" species that are exploited by fisheries operating within EEZs.**

Theories and empirical experience have shown that fisheries targeting resources that straddle between political boundaries complicate fisheries management and potentially reduce its effectiveness to achieve their stated objectives [@Munro:2002uf; @Englander:2019dq].**For instance,** climate-driven changes in species distributions have led to conflicts between nations that were driven, at least partly, by changes in the proportion of captures [@Munro:2002uf], quota allocation, and fishery newcomers [@Spijkers:2017ij]. Moreover, climate change is likely to exacerbate these types of conflict and presents new challenges for political relations between neighboring countries [@Pinsky:2018cb] and for fisheries management [@Miller:2013iv]. Therefore, having an accurate understanding of the distribution and scale of transboundary and straddling fish stocks and fisheries is important to inform their sustainable management, particularly under climate change.

Forty years after the formal adoption of UNCLOS **and the subsequent 1995 United Nations Fish Stock Agreement for the cooperation on the management and conservation of straddling resources [@UnitedNations:1995fs], accurate estimates of the number of exploited marine species that are shared by neighboring nations are still unclear. With limited biogeographical data, an informed guess suggested that there were approximately 500 to 1500 exploited transboundary stocks [@Caddy:1997ue]. Literature review undertaken by a more recent study included 344 shared taxa and the historical contributions to fisheries [@Teh:2015gd].** However, these previous studies did not include species' biogeography in the quantification of transboundary stocks. Here, we aim to estimate the number of exploited marine species shared by neighboring countries and determine their contribution to global and regional catch and fisheries revenue. Moreover, we categorize species according to their catch trends while identifying the differences between species by habitat preference. We hypothesize that the methodological constraints of previous studies has resulted in an underestimation of the number of transboundary stocks and their contribution to global catch and revenue.

We overlaid the known distribution of 938 commercially valuable marine species that are responsible for an average of 96.5% of global EEZ catch reconstructed by the *Sea Around Us* at the species level between 2005 and 2014, and 280 EEZs of 198 coastal **countries (See methods**). We define a ‘stock’ unit as a species in an EEZ, instead of using a population ecology definition (genetically or morphological distinct units), due to a lack of such biological information for almost all the species included in this study [@Teh:2015gd]. While we acknowledge that a species could have multiple stocks within an EEZ, many fisheries within a country or EEZ are managed at the species instead of stock level **(or even as groups containing multiple species). For example, shrimps (e.g. Litopenaeus stylirostris or Farfantepenaeus californiensis) in the Pacific coast of Mexico [@DiarioOficialdelaFederacionDOF:2018wt] and hammerhead shark (*Sphyrna zygaena*) in Peru [@MAP:2017uf] include multiple populations that are exploited by each country’s fisheries but they are managed at the species level in both cases.** Also, recent research has suggested strong connectivity between fish stocks, even when separated by thousands of kilometers [@Ramesh:2019va] through larval dispersal and **adult migration e.g. fishing migrating between reefs in neighboring EEZs [Levin:2018in]. Such ecological connectivity further highlights to the importance of accounting for transboundary stocks using biogeographical information and making the results relevant from a fisheries management perspective**. For this analysis we only considered shared species between neighboring EEZs, rather than the species’ extended distribution (e.g. we did not include the high seas). We rely on multiple data sources including occurrence, distribution models and catch data and only consider a species to be present in a grid cell if all data sources predicted positive occurrence (see Methods).

@@

**We identified 633 transboundary exploited species in the world (67.5% of the 938 species analyzed), almost double the previous estimates [@Teh:2015gd].** Between 2005 and 2014, national fleets targeting these transboundary species within EEZs **caught** an **annual** average of 48.5 million tonnes, representing 82.3% of the EEZ-derived catches reconstructed by the *Sea Around us* at the species level. These catches generated **a yearly average of USD 77,591 million in fishing revenue (78.5% of the global fishing revenue) during the same time period.** These values are more than the previously estimated 34.2 million tonnes in catches and USD 40,187 million (**in 2019 value**) in fishing revenue from shared stocks [@Teh:2015gd].**When we re-estimated the transboundary species’ catches using catch data that are consistent with those used in previous studies (FAO global reported data in year 2006), the 633 transboundary species we identified here accounted for 40.4 million tonnes of catches i.e., 18% higher than previously estimates. Our results suggests that the participation of transboundary species in global catch and fishing revenue might have been previously underestimated due to an incomplete knowledge on the transboundary nature of marine fished species.** The 305 species that are considered non-transboundary (termed as ‘discrete’ species here, see methods - Determining transboundary species trait) contributed a much smaller proportion of catches and revenues compared to those from the 633 transboundary species identified in this study. These discrete species represented only **an annual average of** 2.8 million tonnes and 4,282 million USD between 2005 and 2014. These results underscore the importance of transboundary species at the global level.

In many cases, according to our categorization criteria, a transboundary species can distribute in multiple EEZs but only counted as transboundary in a subset of EEZs (see Methods). For example, the distribution of Peruvian anchoveta (*Engraulis ringens*) spans over the EEZs of Peru, Chile, and Ecuador. However, our study only counted the transboundary stocks in Peru and Chile [@Cashion:2018cg]. Anchoveta in Ecuadorian waters only include a small proportion of the shared distribution range **(spatial threshold between Ecuador and Peru < 10%; see methods - Criteria 3)** and thus do not meet our criteria to be considered as transboundary stock. Situation similar to the example of Peruvian anchoveta in Ecuadorian waters is common amongst the 633 transboundary species identified in this study. Overall, 590 of the 633 transboundary species have stocks in EEZs that are excluded from being counted as transboundary because they do not meet all our categorization criteria. The **annual average contribution from all the excluded stocks of transboundary species to fisheries (e.g., Ecuador’s anchoveta catch) between 2005 and 2014** was 10.8 million tonnes, representing USD 19,243 million in fishing revenue **over the same time period.**

We group countries according to the United Nations sub-regions and found that transboundary species are particularly economically important for Northern America (**average country** revenue = 4,680 $\pm$ 6,000 million USD) and Eastern Asia (3,779 $\pm$ 3,093 million USD). The estimated **per country** revenues from transboundary species in these two regions are significantly different from other regions (one-way ANOVA; *DF*(16,165) = 5.081, *p* < 0.001, $\alpha$ = 0.05) (Fig. 1B). China (USD 7,284 million) and the USA (USD 11,604 million) contribute 55% and 82% to the **annual average** revenues in **Eastern** Asia and **Northern America**, respectively. In addition to China and the USA, Russia (USD 7,379 million), Peru (USD 6,044 **million**) and Japan (USD 3,907 **million**) are among the top five countries with the most fishing revenue generated from transboundary species between 2005 and 2014. These five nations were responsible for 41% of the yearly global fisheries revenue from transboundary species.

@@

![**Figure 1.** Number of transboundary species and their contribution to global fisheries catches and revenue. A) The number of species and amount of revenue are represented by color coding of EEZs and land polygons, respectively. B) Contribution of transboundary species to regional revenue (left) and catch (right). Regions according to the United Nations sub-regions: Northern America (N. America), Southern America (S. America), Latin America and the Caribbean (Ltn. Ame. & the Car.) Points = mean $\pm$ sd. Revenue in 2019 real USD.](../Figures/Fig1.png)

Peru and Russia, having the two largest fisheries by total catch in the world [@FAO:2018KK], caught over 5.8 million tonnes annually by each country on average between 2005 and 2014 (Supplementary Fig. S1). Peru's catches are mainly composed of Peruvian anchoveta (*Engraulis ringens*), shared with Chile, and account for 79% of the **national** transboundary species production. Recently these two countries have signed an agreement to work towards standardized stock assessments through coordinated management of this species [@UNDP:2016wp]. **Similarly, Russia, Japan and the USA have made a management agreement over Alaskan pollock (*Theragra chalcogramma*) in the Bering Sea since 1988 [@NOAA:gX6CtUTk].** Transboundary species also contribute largely to fisheries in **Eastern** Asia (one-way ANOVA; *Df*(16,158) = 2.265, *p* = 0.005, $\alpha$ = 0.05). Particularly, China the world's top fish producer [@FAO:2018KK] **gets** one third (5.1 million tonnes) of the total catches from transboundary species, followed by Japan (1.8 million tonnes) and South Korea (1.06 million tonnes). Differences in transboundary fisheries can also be illustrated in terms of catch-revenue over the area of the EEZ $km^2$ (Fig. 2). As an example, the aggregated area of all EEZs of Northern European countries where transboundary species are present is 3.3 million $Km^2$, the $6^{th}$ smallest of the 17 groups analyzed (Supplementary Table S1). However, it has the second **highest average** revenue (USD 26.1 thousand per $Km^2$) and the **highest average*** catch (19.9 tonnes per $Km^2$) of transboundary species per EEZ area between 2005 and 2014. Countries from Western Europe get significantly more revenue from transboundary fisheries per $Km^2$ than from other regions (one-way ANOVA, *Df*(16,165) = 3.267, *p* < 0.001, $\alpha$ = 0.05; Tukey's post hoc test *p* $\leq$ 0.05).

@@

![Figure 2. Weighted benefits of transboundary species by $km^{2}$ and UN sub-region. A) Units in thousand 2019 USD. B) Units in tonnes. Points are sub-region averages $\pm$ s.d. by country](../Figures/Fig2.png)

**We determined the catch trend of each species within each EEZ (Fig. 3). While previous studies demonstrated that catch trends can be used to infer whether a stock is healthy, re-building, over-exploited or collapsed [@Kleisner:2011wn], several factors can influence stock status. Therefore, our intention here is to examine where the catch trends of transboundary species are different from non-transboundary species between 2005 and 2010 relative to historic catch since 1951 (See methods).** Catch trends are classified as increasing (Category A), constant (Category B) or decreasing (Category C). We found significant differences within all catch trend categories for transboundary species (one-way ANOVA, *Df*(2,459) = 47.94, *p* < 0.001, $\alpha$ = 0.05; Tukey's post hoc test *p* $\leq$ 0.001) as opposite to discrete species where no significant differences where registered (one-way ANOVA, *Df*(2,106) = 1.885, *p* = 0.157; $\alpha$ = 0.05). We also found significant differences in catch trends when directly comparing transboundary to discrete species (MANOVA, *Df*(2,459) = 19.001, *p* < 0.001). **Overall, transboundary species that are only fished by one country are generally less likely to have a decreasing catch trend compared to when the species is shared by neighboring countries (Supplementary Table S2)**.

```{r Call_Circ_Country, eval = F, fig.width=6, fig.height=4, echo=FALSE, fig.pos= "H", fig.cap="Number of transboundary species by catch category. Category A, increasing catch; Category B, constant catch; Category C, decreasing catch"}

@@

![Figure 3. Number of transboundary species by catch trend and EEZ. Category A, Growing trend; Category B, Constant trend; Category C, Shrinking trend](../Figures/Fig3.png)

@@

Transboundary fisheries are abundant and of economic importance for some of the top countries with past fisheries-related conflicts including Canada, the USA, Russia and the European Union (EU) [@Spijkers:2019cz]. For example, since 2007, the EU, Norway, Iceland, and the Faroe Islands have been at odds over the size and relative allocation of total allowable catches for Atlantic mackerel (*Scomber scombrus*) due to a climate-driven distribution shift [@Spijkers:2017ij]. Atlantic Mackerel is a transboundary species that **straddles into the high seas**. Amongst the countries involved in the 2007 Atlantic mackerel fisheries dispute, mackerel contributed an annual average (between 2005 and 2014) of 598.4 thousand tonnes (8.19%) to these countries’ total catches and USD 850.2 million (7.17%) to their total fishing revenue. Climate change is expected to continue changing the distribution and share proportion of fish stocks among countries up to a degree that new transboundary species will emerge [@Pinsky:2018cb], while some species will exit EEZs [@Oremus:2020id]. Exploring the detailed effects of climate change in the distribution of shared species is key to develop local adaptation methods to anticipate negative impacts to sustainability. **For example, understanding how climate change will modify the proportion of transboundary species that each neighboring EEZ share [@PalaciosAbrantes:2020vt], the time-frame at when the changes in sharing will happen, and the economic consequence of such effects can inform the development of more climate-resilient international treaties [@RashidSumaila:tp]**

Over 130 marine fish species occur in tropical and subtropical waters around the world [@Nelson:2016kk], from highly migratory species associated to pelagic-oceanic ecosystem like tunas (*Thunnus sp.*), to less mobile reef-associated species like greater amberjack (*Seriola dumerili*), to species found in demersal ecosystems like gilthead seabream (*Sparus aurata*). Species associated to pelagic-oceanic ecosystems are the only group whose EEZ range (i.e., the number of EEZs where the species occur as transboundary) was significantly different from other groups (one-way ANOVA; *DF*(5,597) = 53.82, *p* < 0.001, $\alpha$ = 0.05; Tukey's post hoc test *p* < 0.05) with a median of 40 EEZs per species. The median for species of all other ecosystem preferences is close to, or less than, 20, as many of these species have a narrower distribution or are less mobile (Fig. 4A). **This result is likely an effect of the broad distribution that many of the species with preference for pelagic-oceanic ecosystems have. Many pelagic-oceanic species are highly migratory and straddle across one EEZ to another while crossing the high seas. Thus, the number of non-neighboring EEZs sharing a highly migratory species can be over 100, like the case of bigeye tuna (Thunnus obesus) (Fig. 4A). Due to their large migration patterns and presence in areas beyond national jurisdiction, many highly migratory species are managed by Regional Fisheries Management Organizations (RFMOs).**

@@

![Figure 4. Number of EEZs shared by transboundary species. A) Number of EEZs by species organized by associated habitat according to FishBase. Showing only species that share > 20 EEZs. B) Total catch between 2005 and 2014 of top five fishing countries capturing the top five shared species of each ecosystem preference (shown in bold on A). Note that there could be > 5 species due to similar sharing values. ** The category "Other" consists of species that have no classification according to FishBase.**](../Figures/Fig4.png)

**Many transboundary species are not considered highly migratory but still shared by numerous neighboring EEZs (e.g. Belone belone). In addition, many fish stocks have meta-populations that are connected through larvae dispersal with ‘source’ populations potentially supporting ‘sink’ populations thousands of kilometers away [@Ramesh:2019va]. For example, despite coral reef-related species being among the less shared neighboring EEZs identified in this study (Fig. 4), coral regions often share multiple species through larval connectivity [@Schill:2015fh] and adult movement [@Levin:2018in]. Species with narrower ranges are often managed at the national level due to localized populations [@Payne:2008vm] but the potential boundary-crossing of these species must not be ignored. In some areas like the Caribbean and the Western Indian Ocean, transboundary marine protected areas have been identified as potential tools for achieving fisheries and conservation goals [@Levin:2018in; @Perez:2019dm]. The proper management of coral reef species is very important for many coastal communities, particularly in developing regions, that are highly dependent on subsistence fishing for food security and livelihoods [@CisnerosMontemayor:2016gq; @Wabnitz:2018gf]. For instance some of the top countries fishing transboundary reef and pelagic-oceanic associated species (Fig. 4B**) are among the highest fish consumption countries [@FAO:2018KK]. In countries like the Philippines both pelagic and reef fishes contribute substantially to both local food security and livelihood [@Cabral:2018hc].

@@

Occurrence data was collected **by Reygondeau (2019)** from five publicly available repositories: FishBase (http://fishbase.org), the Global Biodiversity Information Facility (GBIF; https://www.gbif.org/), the Ocean Biogeographic Information System (OBIS; https://obis.org/), the Intergovernmental Oceanographic Commission (IOC; http://ioc-unesco.org), and the International Union for Conservation of Nature (IUCN; https://www.iucn.org/technical-documents/spatial-data) [@Reygondeau:2019uh].

@@

Categories <- c("A - Increasing catch trend",
                  "B - Constant catch trend",
                  "C - Decreasing catch trend Fishery",
                  "No Status")
